04-analisis

### 2024-05-15 03-methods.qmd
update plots and tables analisis
Primer cambios
Cambios en method.qmd y analisis.qmd

## Methods

@@

The main data source for the analysis is the First Study of Citizenship Education in Chile, carried out by the Education Quality Agency of the Ministry of Education. The application date was November 9, 2017. The target population of this study is Eighth-grade students from 242 schools. In the data, there are 8,589 students and 6,770 parents. This database was analyzed with the R package “ResponsePatterns” to detect possible repetitive and “careless” response patterns and thus contribute to a higher quality of research data (Gottfried, et al 2022). Responses from 171 students and 79 parents were removed, which, when merged, gave a total of 6,270 valid cases.

The analysis of school variables includes data from the Ministry of Education's SIMCE 2017 database. This database contains information at the school level, such as the administrative dependency, its socioeconomic classification, and the achievement scores obtained in the mathematics and language census tests. It is available for free use on the MINEDUC \[web page\].

After eliminating missing cases, the final sample used in the analysis was based on 5,047 students and parents of 231 schools for the dependent variable of access to social services.

@@

**Dependent variables**

This study has three dependent variables related to the justification of social inequality in specific policy domains. The first asks whether access to social services should be conditional on income, i,.e., “It is just that in Chile people who can pay have a better education for their children”. Students rated their preferences using the following responses: “strongly disagree”, “Disagree”, “Agree”, and ”strongly agree”. An average index is built with these items (Cronbach's Alpha = 0,86). (Apéndice: items en español y su correspondiente traducción al inglés)

@tbl-desc-dependientes shows the items used, their response categories, and their frequencies.

```{r echo=FALSE}
pacman::p_load(summarytools, dplyr)
load("input/data/dependientes.RData")

```

@@

**Independent variables**

For the primary independent variable, the perception of meritocracy, five items address the perception of rewards according to effort and intelligence at the school and societal levels. At the school level, students answer whether “Intelligence is important to get good grades” and “Effort is important to get good grades”. At the societal level, students respond to the following questions: “In Chile, people are rewarded for their efforts”, “In Chile, people get what they deserve”, and “In Chile, people are rewarded for their intelligence and skills”. Each item was answered on a four-point scale ranging from “Strongly disagree” to “Strongly agree”.

The rest of the independent variables are divided into individual and school levels. At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home. Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection. @tbl-desc-independent shows the items used, their response categories, and their frequency.

@@

The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees. @tbl-desc-school shows the items used, response categories, and frequency.

@@

The data has a hierarchical structure of students nested in schools, so the model estimation is performed in a multilevel (random effects) framework. This modeling approach lets us correctly estimate individual and contextual effects in a single model. We estimate cumulative link mixed models for the ordinal dependent variables, whereas we use linear mixed effects models for the average index of inequality justification.

>>>>

## Data, variables and method

@@

The main data source corresponds to the First Study of Civic Education in Chile, conducted by the Agency for Quality Education of the Ministry of Education in 2017. This database is based on the administration of a civic knowledge test to students and questionnaires to various members of the educational community. The target population consists of 8th-grade students in 242 schools across the country. Initially, the database comprises 8,589 student observations and 6,770 parent observations. However, considering the survey's unit of analysis and to ensure higher data quality, 171 student cases and 79 parent cases exhibiting repetitive and careless response patterns were removed (Gottfried et al., 2022). Additionally, data from the System for Measuring the Quality of Education (SIMCE) conducted by the Ministry of Education for the year 2017 are employed, providing school-level information such as administrative dependence, socioeconomic classification, and results obtained in standardized mathematics and language tests. After variable processing and removal of missing cases, the final database for this study includes a two-level stratified sample, composed of 5,047 students and parents (level 1) nested within 231 schools (level 2).

@@

***Individual level***

**Justification of Inequality**: The dependent variable in this study is the justification of social inequality in specific policy domains, namely health, education, and pensions. This construct is measured through three variables addressing the degree of justification regarding whether access to social services in these domains should be income conditional. Justification of inequality in health is measured by the item "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes," while in education, it is measured using the statement "Is it fair in Chile that people who can afford it have better education for their children." In the case of pensions, it is measured through the item "Is it fair in Chile that people with higher incomes can have better pensions than those with lower incomes." In all cases, respondents declare their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4). Additionally, we include a summarized indicator of preferences for market justice (REF), measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes).

```{r echo=FALSE}
pacman::p_load(summarytools, dplyr, sjlabelled)
load("input/data/dependientes.RData")
dependientes$bienestar <- set_label(dependientes$bienestar, "Market Justice Preferences")
```

@@

**Perception of Meritocracy**: The main independent variable refers to the perception of meritocracy, operationalized through five items addressing the perception of rewards based on talent and intelligence at both the school and societal levels. At the school level, students respond to whether "Intelligence is important for getting good grades" and "Effort is important for getting good grades". At the societal level, students respond to the following questions: "In Chile, people are rewarded for their effort", "In Chile, people get what they deserve" and "In Chile, people are rewarded for their intelligence and skills". Each item was answered on a four-point Likert scale ranging from "completely disagree" (1) to "completely agree" (4).

The analysis includes several individual control variables that have been identified as influential in recent research. These variables encompass family socioeconomic status, assessed by the highest level of parental education, the number of books in the household, and an index measuring access to technologies, all aimed at mitigating compositional effects (REFS). @tbl-desc-independent shows the independent variables used, their response categories and their frequencies.

@@

***Contextual level***

This study examines four variables at the school level. Firstly, we assess the administrative dependence of schools, categorized as "Public," "Subsidized Private," or "Private." Secondly, we consider the socioeconomic classification of schools, determined by the Ministry of Education and measured on an ordinal scale with five categories ranging from "low" (1) to "high" (5). Thirdly, we evaluate the performance level of schools in SIMCE standardized tests, categorized as "low," "medium," or "high." Lastly, we analyze the proportion of parents with university or postgraduate education level at each school. The items used, response categories, and their frequencies are detailed in @tbl-desc-school.

@@

Given the hierarchical structure of the data, with students nested within schools, model estimation is performed within a multilevel framework. These models are appropriate for capturing both individual and contextual effects in a single analysis, allowing for the estimation of fixed effects between groups and random effects that vary from one group to another [@bell_fixed_2019; @hox_multilevel_2010]. Cumulative link mixed models were employed for the ordinal dependent variables, while linear mixed-effects models were applied for the average market justice preference index.

### 2024-05-15 04-analisis.qmd
Primer cambios
Cambios en method.qmd y analisis.qmd

```{r}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
```

Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X). On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown. For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent. This relationship needs to be clarified in the case of the variables of perception of meritocracy at school. To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society. In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.

### Cumulative link mixed models for the Justification of inequality in education, health and pensions

Three Cumulative link mixed models were estimated for the ordinal dependent variables of justice in differential access to pensions, education, and health according to income. Figure 4 shows the estimation of this regression model containing all the variables used in the study for the three dependent variables separately. However, this figure shows only the effect of meritocracy variables on society and school as independent variables. Complete models are available in the appendix.

```{r}
#| fig-cap: "Odds-ratios of justification of education, pensions and health by social and school meritocracy"
```

Figure 4 shows that for the three dependent variables of justice in differential access to pensions, education, and health, the trend is similar. In the context of school meritocracy, the effects are mixed. As school talent increases, the justification for differentiated access to pensions, education, and health increases; on the contrary, as the school effort variable increases, the justification for differentiated access to pensions, education, and health decreases, keeping the rest of the independent variables constant. Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases.

### Multilevel regression models for market justice preferences

```{r}
#| tbl-cap: "Individual effects of multilevel regression models"
```

Table 4 shows the results of the multilevel estimation for justice market preferences. For this variable, the intraclass correlation obtained shows that the variation between schools corresponds to 4% of the variation of students’ preferences. This means that there is low variance between schools and therefore limits the possibilities of finding effects at the aggregate level.

@@

```{r}
#| tbl-cap: "Contextual effects of multilevel regression models"
```

### Interactions effects

```{r}
#| tbl-cap: "Interactions effects"
```

The interaction terms in Table 6 suggest that students’ family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality. The direction of these effects confirms our initial prediction. Thus, in line with our Hypothesis 3, the relationship between social effort and justification of inequality becomes less positive for those students whose parents achieved university or postgraduate education. That is, lower-status students (measured by parental education) justify more inequality when they adhere more strongly to deservingnessmeritocratic perceptions in Chile. We depict this moderating effect in Figure 1. This result confirms the enlightenment thesis (see above). The same moderating effect is observed for students’ family cultural capital: the relationship between effort (but not deservingness nor social talent) and justification of inequality becomes less positive for students with more than 25 books at home. Interestingly, family cultural capital also moderates the relationship between meritocratic beliefs at school (as measured by students’ belief that effort is important to get good grades) and justification of inequality.

At the school level, we also observe moderating effects of school status and students’ meritocratic beliefs over their justification of inequality. Thus, in line with our Hypothesis 5, high-status schools justify less inequality when, on average, their students have a greater perception of meritocracy in Chile, as measured by the three indicators we used (i.e., effort, deservingness, and talent). Finally, as we stated in Hypothesis 7, lowhigh-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile. In Figure 25 we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality. We did not observe these moderating effects at the school level for students’ meritocratic beliefs in the school (i.e., the idea that effort or talent are important to get good grades).

```{r}
#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
```

>>>>

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4, 
               ordinal,
               sjPlot, 
               texreg,
               kableExtra)

options(scipen=999)
rm(list = ls())

```

```{r data, echo=FALSE}

load("input/data/data_index.RData")
load("input/data/deservingness.RData")
load("input/data/soc_effort.RData")
load("input/data/soc_talent.RData")
load("input/data/school_effort.RData")
load("input/data/school_talent.RData")
load("input/data/data_rec.RData")

```

```{r functions, echo=FALSE}

miles <- function(x) {
  format(round(as.numeric(x),0), big.mark = ".")
}

decimales <- function(x) {
  format(round(as.numeric(x), 2), decimal.mark = ",")
}

# set theme

#windowsFonts(`Times New Roman` = windowsFont("Times New Roman"))

theme_nice <- function() {
  theme_bw() +
    theme(text = element_text(family = "serif"))

}

theme_set(theme_nice())

```

@fig-bivariate shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram in @fig-hypotheses). On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown. For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent. This relationship needs to be clarified in the case of the variables of perception of meritocracy at school. To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society. In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.

```{r echo=FALSE}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
#| fig-asp: 1.22
#| label: fig-bivariate

plot_soc_effort <- soc_effort %>% 
  mutate(justice = str_to_title(justice)) %>% 
  ggplot(aes(x = esfuerzo_soc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_grey() +
  labs(x = "Social effort",
       y = "Market justice preferences") +
  theme(legend.position = "none",
        text = element_text(size = 11))

plot_deservingness <- deservingness %>% 
  mutate(justice = str_to_title(justice)) %>% 
  ggplot(aes(x = merito_soc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_grey() +
  labs(x = "Deservingness",
       y = "Market justice preferences") +
  theme(legend.position = "none",
        text = element_text(size = 11))

plot_soc_talent <- soc_talent %>% 
  mutate(justice = str_to_title(justice)) %>% 
  ggplot(aes(x = inteligencia_soc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_grey() +
  labs(x = "Social talent",
       y = "Market justice preferences") +
  theme(legend.position = "none",
        text = element_text(size = 11))

plot_school_effort <- school_effort %>% 
  mutate(justice = str_to_title(justice)) %>% 
  ggplot(aes(x = esfuerzo_esc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_grey() +
  labs(x = "School effort",
       y = "") +
  theme(legend.position = "none",
        text = element_text(size = 11))

plot_school_talent <- school_talent %>% 
  mutate(justice = str_to_title(justice)) %>% 
  ggplot(aes(x = inteligencia_esc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_grey() +
  labs(x = "School talent",
       y = "",
       color = NULL,
       shape = NULL) +
  theme(text = element_text(size = 11))

leg <- get_legend(plot_school_talent)

plot_school_talent <- plot_school_talent + theme(legend.position = "none")

ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3)
```

### Multivariate results

The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @fig-odds. This figure illustrates the effect of meritocratic variables on society and school as independent variables; however, the models incorporate all variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).

```{r echo=FALSE}

data_pension <- data_rec %>% dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
reg5_pension <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_pension)
data_salud <- data_rec %>% dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
reg5_salud <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_salud)
data_educ <- data_rec %>% dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
reg5_educ <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_educ)
```

```{r echo=FALSE}
#| fig-cap: "Odds-ratios of justification of pensions, education and health by social and school meritocracy"
#| label: fig-odds
#| fig-asp: 0.8

plot_models(reg5_salud, reg5_educ, reg5_pension,  
            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part. subvencionado", "cod_depe2Part. privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 1.3, #tamaño circulos
            line.size = 1, #tamaño CI
            value.size = 3.5, #tamaño valor efectoss
            spacing = .7, #espacio entre efectos
            grid = TRUE, # separa en un plot por modelo
            vline.color = "firebrick", # linea roja en punto neutro (1)
            axis.labels = rev(c("School talent", 
                            "School effort", 
                            "Social talent",
                            "Social effort", 
                            "Deservingness"
                            )), #con rev porque automatico los tira en otro orden
            show.legend = FALSE,
            m.labels = c("Health", "Education","Pension"),
            colors = gray(level = 0.2))  # variables dependientes

```

@fig-odds demonstrates a similar trend for the three dependent variables of justice in differential access to pensions, education, and health, with all effects being statistically significant. In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant. Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases. Overall, these results provide partial support for our first hypothesis (H1a), which posited that students who perceive higher levels of meritocracy in school would be more likely to justify differential access to pensions, health, and education. This is evidenced by the positive effect of perceiving talent as important for achieving good grades. However, the results strongly support our second hypothesis (H1b), as a higher perception of societal meritocracy among students tends to increase the justification for market-based justice in these social services.

Contextual effects for ordinal models.

```{r echo=FALSE, include=FALSE}

reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data=data_bienestar)
reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_bienestar)
reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_bienestar)
reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)
reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_bienestar)

test01<- anova(reg0_bienestar,reg1_bienestar,test = "Chisq")
test02<- anova(reg0_bienestar,reg2_bienestar,test = "Chisq")
test03<- anova(reg0_bienestar,reg3_bienestar,test = "Chisq")
test04<- anova(reg0_bienestar,reg4_bienestar,test = "Chisq")

test.pvalues1<- test01$`Pr(>Chisq)`[2]
test.pvalues2<- test02$`Pr(>Chisq)`[2]
test.pvalues3<- test03$`Pr(>Chisq)`[2]
test.pvalues4<- test04$`Pr(>Chisq)`[2]

deviance1 <- test01$deviance[2]
deviance2 <- test02$deviance[2]
deviance3 <- test03$deviance[2]
deviance4 <- test04$deviance[2]
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Multilevel models for meritocracy perception and market justice preferences"
#| label: tbl-individual-reg

texreg::htmlreg(list(reg1_bienestar, reg2_bienestar),
        custom.model.names = c("Model 1",
                               "Model 2"),
        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
        omit.coef = "(internet_rec)",
        caption = "",
        custom.coef.names = c("Intercept",  
                              "School talent",
                              "School effort", 
                              "Social talent",  
                              "Social effort",
                              "Deservingness",
                              "Secondary", 
                              "Higher tec.", 
                              "University or posgraduate", 
                              "Missing", 
                              "More than 25 books (Ref. Less than 25)", 
                              "Technology access"),
        groups = list("Parental education (Ref.= 8th grade or less)"=7:10),
        custom.gof.rows=list("Deviance" = c(deviance1,
                                            deviance2),
                             "Deviance Test (p)" = c(test.pvalues1,
                                                       test.pvalues2))
        )

```

Regarding our measure of market justice preferences index, the results of the multilevel linear estimation are shown in @tbl-individual-reg. The intraclass correlation obtained for the null model (not shown) it's equal to 4% (ICC based on [@hox_multilevel_2010, p.15]), which indicates that only a small percentage of the total variance of the students' market justice preferences are associated with the characteristics of schools, and therefore limits the possibilities of finding effects at this second level of analysis.

@@

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Contextual effects of multilevel regression models"
#| label: tbl-contextual-reg

texreg::htmlreg(list(reg3_bienestar, reg4_bienestar),
        custom.model.names = c("Model 1",
                               "Model 2"),
        custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
        omit.coef = c("inteligencia_esc|esfuerzo_esc|inteligencia_soc|esfuerzo_soc|merito_soc|educacion_rec|libros_rec|acc_tec|internet_rec"),
        caption = "",
        custom.coef.names = c("Intercept", 
                              "Prop. university level at school",
                              "Subsidized private", 
                              "Private", 
                              "SES Medium low", 
                              "SES Medium", 
                              "SES Medium high", 
                              "SES High", 
                              "Simce Medium", 
                              "Simce High"),
        groups = list("Administrative dependence (Ref.= Public)"=3:4,
                      "Socioeconomic level (Ref.= Low)"=5:8,
                      "Achievement score (Ref.= Low)"=9:10),
        custom.gof.rows=list("Deviance" = c(deviance3,
                                            deviance4),
                             "Deviance Test (p)" = c(test.pvalues3,
                                                     test.pvalues4))
        )

```

```{r echo=FALSE}
#| tbl-cap: "Interactions effects"
#| label: tbl-interact

pacman::p_load(kableExtra)
load("input/data/interact.RData")
kbl(collapse_rows,
    align = "lccccc", booktabs = T,
    col.names =  c("Variable", 
                   "School talent", 
                   "School effort",
                   "Social talent",
                   "Social effort", 
                   "Deservingness")) %>%
  add_header_above(c(" " = 1,
                     "Market Justice Preferences"=5), 
                   bold = T)%>%  
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                full_width = F) %>%
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(column = 2:6,width = "1.5cm") %>%
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  footnote(general = "***p < 0.001, **p < 0.01, *p < 0.05",
           footnote_as_chunk = T,general_title = "Note: ")

```

The interaction terms in @tbl-interact suggest that students' family background moderates the relationship between their meritocratic perceptions in Chile and their justification of inequality. The direction of these effects confirms our initial prediction. Thus, in line with our Hypothesis 3, the relationship between social effort and justification of inequality becomes less positive for those students whose parents achieved university or postgraduate education. That is, lower-status students (measured by parental education) justify more inequality when they adhere more strongly to deservingnessmeritocratic perceptions in Chile. This result confirms the enlightenment thesis (see above). The same moderating effect is observed for students' family cultural capital: the relationship between effort (but not deservingness nor social talent) and justification of inequality becomes less positive for students with more than 25 books at home. Interestingly, family cultural capital also moderates the relationship between meritocratic beliefs at school (as measured by students' belief that effort is important to get good grades) and justification of inequality.

At the school level, we also observe moderating effects of school status and students' meritocratic beliefs over their justification of inequality. Thus, in line with our Hypothesis 5, high-status schools justify less inequality when, on average, their students have a greater perception of meritocracy in Chile, as measured by the three indicators we used (i.e., effort, deservingness, and talent). Finally, as we stated in Hypothesis 7, low-achieving schools justify more inequality when their students have, on average, stronger meritocratic beliefs in Chile. In @fig-interaction we depict this moderating effect of school achievement for the relationship between deservingness and justification of inequality. We did not observe these moderating effects at the school level for students' meritocratic beliefs in the school (i.e., the idea that effort or talent are important to get good grades).

```{r echo=FALSE, include=FALSE}
load("input/data/data_rec.RData")
data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
                                                 educacion_rec=="Ed Media"~"Low level",
                                                 educacion_rec=="Ed Tecnica"~"Low level",
                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
                                                 educacion_rec=="Ns/Nr"~"Missing"
                                                 ))
data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))

data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
                                                                       cod_grupo=="Medio bajo"~"Bajo",
                                                                       cod_grupo=="Medio"~"Bajo",
                                                                       cod_grupo=="Medio alto"~"Alto",
                                                                       cod_grupo=="Alto"~"Alto"))
data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))

data_bienestar <- na.omit(data_rec)
#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
#                                            educ_rec = 5-just_educ,
#                                            salud_rec = 5-just_salud)
data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)

reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)

plot_interact<- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
  ylim(1,4)+
  labs(x = "Deservingness", y="Market Justice Preferences", colour = "School Simce")+
  theme_bw()
```

```{r echo=FALSE, warning=FALSE}
#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
#| label: fig-interaction

plot_interact

```
